Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-22 paper/graph.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
generated graphs with power-law degree distributions.  They are
commonly used to represent web and social networks.  The user
specifies six parameters that define the graph: the number of
vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
that sum to 1.0 and are discussed below.  The algorithm in
Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
unique non-zero entries in a sparse $N \times N$ matrix $G$, where
each entry $G_{ij}$ represents an edge between graph vertices
$(V_i,V_j)$, with $V_i$ an integer from $1$ to $N$.

@@

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate $G$ \\
\> Reduce $G$: \> Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\

  \end{tabbing}
 \end{minipage}}\end{center}

@@

In the {\it map()}, each of $P$ processors generates a $1/P$ fraction
of the desired edges.  A single random edge $(V_i,V_j)$ is computed
recursively as follows.  Pick a random quadrant of the $G$ matrix with
relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
quadrant as a sub-matrix and select a random quadrant within it, in
the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
the end of the recursion, the final ``quadrant'' is a single non-zero
matrix element $G_{ij}$.

Though $G$ is very sparse, the {\it map()} typically generates some
small number of duplicate edges.  The {\it collate()} and {\it
reduce()} remove the duplicates.  The entire map-collate-reduce
sequence is repeated until the number of resulting key/value (KV)
pairs $N_{kv} = M$.  For reasonably sparse graphs this typically takes
only a few iterations.

@@

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\> MapReduce object $G_{new}$ = new edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Aggregate $G_{new}$ \\
\> Add $G_{new}$ to $G$, $G_{new}$ = empty \\
\> Compress $G$: \> Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\

  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
\> Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = NULL \\
\> Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
\> MapReduce object $y$ = vector, initially empty \\
\> User-defined parameter $\alpha$ \\
\\
while $({residual} > {tolerance})$ \\
\> 1 Compute adjustment for random jumps and zero-outdegree vertices \\
\> \> Add $x$ to $MT$ \\
$*$\> \> Convert $MT$:  Key is vertex $i$ \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$, MultiValue = $(x_i, NULL)$ if $i \in {MT}$; else MultiValue = $x_i$ \\
\> \> \> If (${nvalues} = 2$) $c_{dot} = c_{dot} + x_i$; else $c_{dot} = c_{dot} + \alpha$ \\
\> \> \> Output: Key = $i$, Value = NULL if ${nvalues} = 2$ \\
\> \> $c_{dot}$ = MPI\_Allreduce($c_{dot}$, MPI\_SUM) / $|V|$ \\
\> 2 Compute $y = A^T x$  \\
\> \> Copy $x$ to $y$ \\
\> \> Add $A$ to $y$ \\
$*$\> \> Convert $y$:  Key is row index of $A$ \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$, MultiValue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
\> \> \>  Output: Key = $j$, Value = $A_{ij} y_i$ \\
\> \> Collate $y$:  Key is column index of $A$ \\
\> \> Reduce $y$: \\
\> \> \> Input: Key = $j$, MultiValue = $A_{ij} y_i$ for nonzeroes $A_{ij}$ in column $j$ of $A$ \\
\> \> \> Output: Key = $j$, Value = $\sum_i A_{ij} y_i$ \\
\> 3 Add $c_{dot}$ to $y$; compute $y_{max}$ = max element in $y$ \\
\> \> Map $y$:  \\
\> \> \> Input: Key = $i$, Value = $y_i$ \\
\> \> \> $y_i = y_i + c_{dot}$; $y_{max} = {\rm max}(y_i,y_{max})$ \\
\> \> \> Output:  Key = $i$, Value = $y_i$ \\
\> \> $y_{max}$ = MPI\_Allreduce($y_{max}$, MPI\_MAX) \\
\> 4 Scale $y$ by $y_{max}$ \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i / y_{max}$ \\
\> 5 Compute residual \\
\> \> Add $y$ to $x$ \\
$*$\> \> Convert $x$:  Key is vertex $i$ \\
\> \> Reduce $x$:  \\
\> \> \> Input: Key = $i$, MultiValue = $x_i, y_i$ \\
\> \> \> $residual = {\rm max}(|x_i - y_i|,residual)$ \\
\> \> $residual$ = MPI\_Allreduce($residual$, MPI\_MAX) \\
\> 6 Copy $y$ to $x$ \\
\\
Output: \\
\> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for PageRank vertex ranking, using 4
 MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
 of the MapReduce objects allows several communication intensive {\it
 collate()} operations to be replaced by local {\it convert()}
 operations (marked with asterisks).}

@@

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
triangles, assuming an input graph $G$ of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
matrix.  This exposition follows the triangle-finding algorithm
presented in~\cite{Cohen09}.

@@

Input: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
\> MapReduce object $G_0$ = edges of graph, initially empty \\
\\
1 Copy $G$ to $G_0$ \\
1 Map $G$: \> \> Convert edges to vertices \\
           \> \> \> Input: Key = $(V_i,V_j)$, Value = NULL \\
           \> \> \> Output: Key = $V_i$, Value = $V_j$ \\
           \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
1 Collate $G$ \\
1 Reduce $G$: \> \> Add first degree to one vertex in edge \\
              \> \> \> Input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> \> for each V in MultiValue: \\
              \> \> \> \> if $V_i < V$: Output: Key = $(V_i,V)$, Value = $(D_i,0)$ \\
              \> \> \> \> else: Output: Key = $(V,V_i)$, Value = $(0,D_i)$ \\
2 Collate $G$ \\
2 Reduce $G$: \> \> Add second degree to other vertex in edge \\
              \> \> \> Input: Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
              \> \> \> Output: Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
3 Map $G$: \> \> Low degree vertex emits edges \\
           \> \> \> if $D_i < D_j$: Output: Key = $V_i$, Value = $V_j$ \\
           \> \> \> else if $D_j < D_i$: Output: Key = $V_j$, Value = $V_i$ \\
           \> \> \> else: Output: Key = $V_i$, Value = $V_j$ \\
3 Collate $G$ \\
3 Reduce $G$: \> \> Emit angles of each vertex \\
              \> \> \> Input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> \> for each $V_1$ in MultiValue: \\
	      \> \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
	      \> \> \> \> \> if $V_1 < V_2$: Output: Key = $(V_1,V_2)$, Value = $V_i$ \\
	      \> \> \> \> \> else: Output: Key = $(V_2,V_1)$, Value = $V_i$ \\
4 Add $G_0$ to $G$ \\
4 Collate $G$ \\
4 Reduce $G$: \> \> Emit triangles \\
              \> \> \> Input: Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
              \> \> \> if NULL exists in MultiValue: \\
	      \> \> \> \> for each non-NULL V in MultiValue: \\
	      \> \> \> \> \> Output: Key = $(V_i,V_j,V)$, Value = NULL \\
\\
Output: \\
\> MapReduce object $G$ = triangles in graph: Key = $(V_i,V_j,V_k)$, Value = NULL \\

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for triangle enumeration, using 2 MapReduce objects $G$ and $G_0$}

@@

A connected component of a graph is a set of vertices where all pairs
of vertices in the set are connected by a path of edges.  A sparse
graph may contain many such components.  Figure~\ref{fig:cc} outlines
a MapReduce algorithm that labels each vertex in a graph with a
component ID.  All vertices in the same component are labeled with
the same ID, which is the ID of some vertex in the component.  We
assume an input graph $E$ of undirected edges $(V_i,V_j)$.  This
exposition also follows the connected-component algorithm presented
in~\cite{Cohen09}, with the addition of logic that load-balances data
across processors, which is important when one or a few giant
components exist in the graph.

@@

Input: \\
\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
\> MapReduce object $W$ = workspace
\\
Iterate: \\
\> 1 Map $W$ using $E$ as input: Convert edges to vertices \\
    \> \> Input: Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
    \> \> Output: Key = $V_i$, Value = $E_{ij}$ \\
    \> \> Output: Key = $V_j$, Value = $E_{ij}$ \\
\> 1 Add $V$ to $W$ \\
\> 1 Collate $W$: Vertex as key \\
\> 1 Reduce $W$: Emit edges of each vertex with zone of vertex \\
    \> \> Input: Key = $V_i$, MultiValue = $E E E E ... Z$ \\
    \> \> for each $E$ in MultiValue: \\
    \> \> \> Output: Key = $E_{ij}$, Value = $Z_i$ \\
\> 2 Collate $W$: Edge as key \\
\> 2 Reduce $W$: Emit zone re-assignments \\
       \> \> Input: Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
       \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
       \> \> if $Z_i$ and $Z_j$ are different: \\
       \> \> \> Output: Key = $Z_{loser}$, Value = $Z_{winner}$ \\
\> 2 Exit: if $W$ is empty (no output by Reduce 2) \\
\> 3 Map $V$: Invert vertex/zone pairs \\
    \> \> Input: Key = $V_i$, Value = $Z_i$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $V_i$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $V_i$ for a random processor \\
\> 4 Map $V$ using $W$ as input: Add zone reassignments in $W$ to $V$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
\> 4 Collate $V$: Zone ID as key \\
\> 4 Reduce $V$: Emit new zone assignment of each vertex \\
       \> \> Input: Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
       \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
       \> \> partition $Z_{new}$ if number of $V >$ threshold \\
       \> \> for each $V$ in MultiValue: \\
       \> \> \> Output: Key = $V_i$, Value = $Z_{new}$ \\
\\
Output: \\
\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\

  \end{tabbing}
 \end{minipage}}\end{center}

@@

The third {\it map()} inverts the vertex/zone KV pairs to become
zone/vertex KV pairs.  The partitioning logic described in
Figure~\ref{fig:cc} for this step (and subsequent ones) is described
below.  The fourth {\it map()} adds the changed zone assignments in
$W$ to the set of KV pairs in $V$.  The fourth {\it collate()} can now
collect all the vertices in each zone along with any reassignments of
that zone ID.  Since a zone could collide with multiple other zones on
the same iteration due to shared edges, the new zone ID for a zone
becomes the minimum ID of any of its neighboring zones.  If no zone
reassignment values appear in the multivalue resulting from the fourth
{\it collate()}, the zone ID is unchanged.  The final {\it reduce()}
emits a KV pair for each vertex in the zone, with the vertex as a key
and the new zone ID as a value.  This is an updated version of
MapReduce object $V$, ready to be used as input for the next
iteration.

@@

During the third {\it map()}, if the zone has been partitioned, the
vertex is assigned to a random processor and the processor ID bits are
added to the zone ID, as indicated by the $Z_i^+$ notation in
Figure~\ref{fig:cc}.  Likewise, if $Z_i$ has been partitioned, the
fourth {\it map()} emits the KV pair for the zone ID as
($Z_i^+$,$Z_{winner}$) instead of as ($Z_i$,$Z_{winner}$).  In this
case ($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once
for each processor, and once as if $Z_i$ had not been partitioned (for
a reason discussed below).

@@

At each iteration, the current set of edges in $E$ is examined to
identify winning vertices.  Vertices adjacent to winners are flagged
as losers, and all edges which contain a loser vertex are marked for
removal at the start of the next iteration.  Vertices with all their
edges marked as losers are also flagged as winners.  At each iteration
marked edges are removed from the graph, and winning vertices are
added to the MIS.  When the graph is empty, the MIS is complete.

@@

In the second {\it reduce()}, if $V_i$ won the comparison with all its
neighbor vertices, then it becomes a ``winner'' vertex.  All vertices
connected to the winner are emitted with $V_i$ tagged with a winner
flag.  If $V_i$ is not an overall winner, its neighbor vertices are
emitted without the winner flag on $V_i$..  The second {\it collate()}
collects this new information for each vertex.

@@

For a distributed graph, global information about edge adjacencies and
labeled vertices is not available.  Instead, following the
Bellman-Ford-style algorithm of~\cite{SSSPMapReduce, Bellman58,
Ford62}, every edge of the graph is visited in each iteration and
vertex distance are updated.  A MapReduce formulation of this
algorithm, described in Figure~\ref{fig:sssp}, begins with a graph $E$
of directed edges $(V_i,V_j)$.  As generated by the R-MAT procedure
discussed earlier, this may include self-edges $(V_i,V_i)$ and loop
edges (both $(V_i,V_j)$ and $(V_j,V_i)$).  These are handled as
special cases within the algorithm to avoid spurious contributions to
the shortest path distances.  MapReduce object $V$ stores the distance
of each vertex $V_i$ from the source vertex $V_s$.

>>>>

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT matrices~\cite{RMAT} are recursively
generated graphs with power-law degree distributions.  They are
commonly used to represent web and social networks.  The user
specifies six parameters that define the graph: the number of
vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
that sum to 1.0 and are discussed below.  The algorithm in
Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
unique non-zero entries in a sparse $N \times N$ matrix $G$, where
each entry $G_{ij}$ represents an edge between graph vertices
$(V_i,V_j)$, with $V_i$ an integer from $1$ to $N$.

@@

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate $G$ \\
\> Reduce $G$: \> Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

In the {\it map()}, each of $P$ processors generates a $1/P$ fraction
of the desired edges.  A single random edge $(V_i,V_j)$ is computed
recursively as follows.  Pick a random quadrant of the $G$ matrix with
relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
quadrant as a sub-matrix and select a random quadrant within it, in
the same manner.  Perform this process a total of $n$ times where $N = 2^n$.  At
the end of the recursion, the final ``quadrant'' is a single non-zero
matrix element $G_{ij}$.

Though $G$ is very sparse, the {\it map()} typically generates some
small number of duplicate edges.  The {\it collate()} and {\it
reduce()} remove the duplicates.  The entire map-collate-reduce
sequence is repeated until the number of resulting key/value (KV)
pairs $N_{kv} = M$.  For reasonably sparse graphs,
only a few iterations are typically needed.

@@

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\> MapReduce object $G_{new}$ = new edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Aggregate $G_{new}$ \\
\> Add $G_{new}$ to $G$, $G_{new}$ = empty \\
\> Compress $G$: \> Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
\> Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = NULL \\
\> Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
\> MapReduce object $y$ = vector, initially empty \\
\> User-defined parameter $\alpha$ \\
\\
while $({residual} > {tolerance})$ \\
\> 1 Compute adjustment for random jumps and zero-outdegree vertices \\
\> \> Add $x$ to $MT$ \\
$*$\> \> Convert $MT$:  Key is vertex $i$ \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$, MultiValue = $(x_i, NULL)$ if $i \in {MT}$; else MultiValue = $x_i$ \\
\> \> \> If (${nvalues} = 2$) $c_{dot} = c_{dot} + x_i$; else $c_{dot} = c_{dot} + \alpha$ \\
\> \> \> Output: Key = $i$, Value = NULL if ${nvalues} = 2$ \\
\> \> $c_{dot}$ = MPI\_Allreduce($c_{dot}$, MPI\_SUM) / $|V|$ \\
\> 2 Compute $y = A^T x$  \\
\> \> Copy $x$ to $y$ \\
\> \> Add $A$ to $y$ \\
$*$\> \> Convert $y$:  Key is row index of $A$ \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$, MultiValue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
\> \> \>  Output: Key = $j$, Value = $A_{ij} y_i$ \\
\> \> Collate $y$:  Key is column index of $A$ \\
\> \> Reduce $y$: \\
\> \> \> Input: Key = $j$, MultiValue = $A_{ij} y_i$ for nonzeroes $A_{ij}$ in column $j$ of $A$ \\
\> \> \> Output: Key = $j$, Value = $\sum_i A_{ij} y_i$ \\
\> 3 Add $c_{dot}$ to $y$; compute $y_{max}$ = max element in $y$ \\
\> \> Map $y$:  \\
\> \> \> Input: Key = $i$, Value = $y_i$ \\
\> \> \> $y_i = y_i + c_{dot}$; $y_{max} = {\rm max}(y_i,y_{max})$ \\
\> \> \> Output:  Key = $i$, Value = $y_i$ \\
\> \> $y_{max}$ = MPI\_Allreduce($y_{max}$, MPI\_MAX) \\
\> 4 Scale $y$ by $y_{max}$ \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i / y_{max}$ \\
\> 5 Compute residual \\
\> \> Add $y$ to $x$ \\
$*$\> \> Convert $x$:  Key is vertex $i$ \\
\> \> Reduce $x$:  \\
\> \> \> Input: Key = $i$, MultiValue = $x_i, y_i$ \\
\> \> \> $residual = {\rm max}(|x_i - y_i|,residual)$ \\
\> \> $residual$ = MPI\_Allreduce($residual$, MPI\_MAX) \\
\> 6 Copy $y$ to $x$ \\
\\
Output: \\
\> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank 
  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for PageRank vertex ranking, using four
 MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
 of the MapReduce objects allows several communication intensive {\it
 collate()} operations to be replaced by local {\it convert()}
 operations (marked with asterisks).}

@@

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
triangles, assuming an input graph $G$ of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e., an upper-triangular R-MAT
matrix.  This exposition follows the triangle-finding algorithm
presented in~\cite{Cohen09}.

@@

Input: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
\> MapReduce object $G_0$ = edges of graph, initially empty \\
\\
1 Copy $G$ to $G_0$ \\
1 Map $G$: \> \> Convert edges to vertices \\
           \> \> \> Input: Key = $(V_i,V_j)$, Value = NULL \\
           \> \> \> Output: Key = $V_i$, Value = $V_j$ \\
           \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
1 Collate $G$ \\
1 Reduce $G$: \> \> Add first degree to one vertex in edge \\
              \> \> \> Input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> \> for each V in MultiValue: \\
              \> \> \> \> if $V_i < V$: Output: Key = $(V_i,V)$, Value = $(D_i,0)$ \\
              \> \> \> \> else: Output: Key = $(V,V_i)$, Value = $(0,D_i)$ \\
2 Collate $G$ \\
2 Reduce $G$: \> \> Add second degree to other vertex in edge \\
              \> \> \> Input: Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
              \> \> \> Output: Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
3 Map $G$: \> \> Low degree vertex emits edges \\
           \> \> \> if $D_i < D_j$: Output: Key = $V_i$, Value = $V_j$ \\
           \> \> \> else if $D_j < D_i$: Output: Key = $V_j$, Value = $V_i$ \\
           \> \> \> else: Output: Key = $V_i$, Value = $V_j$ \\
3 Collate $G$ \\
3 Reduce $G$: \> \> Emit angles of each vertex \\
              \> \> \> Input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> \> for each $V_1$ in MultiValue: \\
	      \> \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
	      \> \> \> \> \> if $V_1 < V_2$: Output: Key = $(V_1,V_2)$, Value = $V_i$ \\
	      \> \> \> \> \> else: Output: Key = $(V_2,V_1)$, Value = $V_i$ \\
4 Add $G_0$ to $G$ \\
4 Collate $G$ \\
4 Reduce $G$: \> \> Emit triangles \\
              \> \> \> Input: Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
              \> \> \> if NULL exists in MultiValue: \\
	      \> \> \> \> for each non-NULL V in MultiValue: \\
	      \> \> \> \> \> Output: Key = $(V_i,V_j,V)$, Value = NULL \\
\\
Output: \\
\> MapReduce object $G$ = triangles in graph: Key = $(V_i,V_j,V_k)$, Value = NULL 
  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for triangle enumeration, using two MapReduce objects $G$ and $G_0$}

@@

A connected component of a graph is a set of vertices where all pairs
of vertices in the set are connected by a path of edges.  A sparse
graph may contain many such components.  Figure~\ref{fig:cc} outlines
a MapReduce algorithm that labels each vertex in a graph with a
component ID.  All vertices in the same component are labeled with
the same ID, which is the ID of some vertex in the component.  We
assume an input graph $E$ of undirected edges $(V_i,V_j)$.  This
exposition follows the connected-component algorithm presented
in~\cite{Cohen09}, with the addition of logic that load-balances data
across processors, which is important when one or a few giant
components exist in the graph.

@@

Input: \\
\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
\> MapReduce object $W$ = workspace
\\
Iterate: \\
\> 1 Map $W$ using $E$ as input: Convert edges to vertices \\
    \> \> Input: Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
    \> \> Output: Key = $V_i$, Value = $E_{ij}$ \\
    \> \> Output: Key = $V_j$, Value = $E_{ij}$ \\
\> 1 Add $V$ to $W$ \\
\> 1 Collate $W$: Vertex as key \\
\> 1 Reduce $W$: Emit edges of each vertex with zone of vertex \\
    \> \> Input: Key = $V_i$, MultiValue = $E E E E ... Z$ \\
    \> \> for each $E$ in MultiValue: \\
    \> \> \> Output: Key = $E_{ij}$, Value = $Z_i$ \\
\> 2 Collate $W$: Edge as key \\
\> 2 Reduce $W$: Emit zone re-assignments \\
       \> \> Input: Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
       \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
       \> \> if $Z_i$ and $Z_j$ are different: \\
       \> \> \> Output: Key = $Z_{loser}$, Value = $Z_{winner}$ \\
\> 2 Exit: if $W$ is empty (no output by Reduce 2) \\
\> 3 Map $V$: Invert vertex/zone pairs \\
    \> \> Input: Key = $V_i$, Value = $Z_i$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $V_i$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $V_i$ for a random processor \\
\> 4 Map $V$ using $W$ as input: Add zone reassignments in $W$ to $V$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
\> 4 Collate $V$: Zone ID as key \\
\> 4 Reduce $V$: Emit new zone assignment of each vertex \\
       \> \> Input: Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
       \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
       \> \> partition $Z_{new}$ if number of $V >$ threshold \\
       \> \> for each $V$ in MultiValue: \\
       \> \> \> Output: Key = $V_i$, Value = $Z_{new}$ \\
\\
Output: \\
\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

The third {\it map()} inverts the vertex/zone KV pairs to become
zone/vertex KV pairs.  The partitioning logic in
Figure~\ref{fig:cc} for this step (and subsequent ones) is described
below.  The fourth {\it map()} adds the changed zone assignments in
$W$ to the set of KV pairs in $V$.  The fourth {\it collate()} can now
collect all the vertices in each zone along with any reassignments of
that zone ID.  Since a zone could collide with multiple other zones on
the same iteration due to shared edges, the new zone ID for a zone
becomes the minimum ID of any of its neighboring zones.  If no zone
reassignment values appear in the multivalue resulting from the fourth
{\it collate()}, the zone ID is unchanged.  The final {\it reduce()}
emits a KV pair for each vertex in the zone, with the vertex as a key
and the new zone ID as a value.  This is an updated version of
MapReduce object $V$, ready to be used as input for the next
iteration.

@@

During the third {\it map()}, if the zone has been partitioned, the
vertex is assigned to a random processor and the processor ID bits are
added to the zone ID, as indicated by the $Z_i^+$ notation in
Figure~\ref{fig:cc}.  Likewise, if $Z_i$ has been partitioned, the
fourth {\it map()} emits the KV pair for the zone ID as
($Z_i^+$,$Z_{winner}$) instead of as ($Z_i$,$Z_{winner}$).  In this
case, ($Z_i$,$Z_{winner}$) is emitted not once, but $P+1$ times, once
for each processor, and once as if $Z_i$ had not been partitioned (for
a reason discussed below).

@@

At each iteration, the current set of edges in $E$ is examined to
identify winning vertices.  Vertices adjacent to winners are flagged
as losers, and all edges that contain a loser vertex are marked for
removal at the start of the next iteration.  Vertices with all their
edges marked as losers are also flagged as winners.  At each iteration
marked edges are removed from the graph, and winning vertices are
added to the MIS.  When the graph is empty, the MIS is complete.

@@

In the second {\it reduce()}, if $V_i$ won the comparison with all its
neighbor vertices, then it becomes a ``winner'' vertex.  All vertices
connected to the winner are emitted with $V_i$ tagged with a winner
flag.  If $V_i$ is not an overall winner, its neighbor vertices are
emitted without the winner flag on $V_i$.  The second {\it collate()}
collects this new information for each vertex.

@@

For a distributed graph, global information about edge adjacencies and
labeled vertices is not available.  Instead, following the
Bellman-Ford-style algorithm of~\cite{SSSPMapReduce, Bellman58,
Ford62}, every edge of the graph is visited in each iteration and
vertex distances are updated.  A MapReduce formulation of this
algorithm, described in Figure~\ref{fig:sssp}, begins with a graph $E$
of directed edges $(V_i,V_j)$.  As generated by the R-MAT procedure
discussed earlier, this may include self-edges $(V_i,V_i)$ and loop
edges (both $(V_i,V_j)$ and $(V_j,V_i)$).  These are handled as
special cases within the algorithm to avoid spurious contributions to
the shortest path distances.  MapReduce object $V$ stores the distance
of each vertex $V_i$ from the source vertex $V_s$.
